---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>"
)
```

# `independencepvalue`

## What does `independencepvalue` do?
`independencepvalue` is an `R` package that tests the independence between two groups of Gaussian variables, where the groups were obtained by thresholding the correlation matrix.  For a brief overview explaining why `independencepvalue` is needed and how it works, see [here](articles/Overview.html).  For the full explanation, you can read our paper [Inferring independent sets of Gaussian variables after thresholding correlations](XXXX.arxiv.org).

## How do I get `independencepvalue`?

To install `independencepvalue`, run the following:

```{r setup, eval=FALSE}
remotes::install_github("ArkajyotiSaha/independencepvalue-project",
                        subdir = "independencepvalue")
```

(And if necessary, first install the `remotes` package by using `install.packages("remotes")`.)

## How do I use `independencepvalue`?

To get started with `independencepvalue`, we recommend trying [this tutorial](articles/Tutorial.html).

## Other information

This R package is created using literate programming with the [litr](https://github.com/jacobbien/litr-project/tree/main/litr) R package. To know more about `independencepvalue` generation, please visit the [package repository](https://github.com/ArkajyotiSaha/independencepvalue-project).